\section{Discussion and bottleneck analysis}

This is where the project earns its title. A roofline plot on its own only shows
that a kernel is slow; it does not say why. Every claim below is tied to a
specific Nsight Compute counter from the table in the results section, measured
on this card, and where a counter contradicted what I expected I have said so
rather than quietly reporting the version that sounded better.

\subsection{SAXPY anchors the memory bound edge}

SAXPY at 16.8 million elements moves 177.9\,MB of real DRAM traffic in
\SI{0.293}{\milli\second}, which is \SI{607.2}{GB/s}. The measured copy ceiling
on this card is \SI{604.1}{GB/s}. SAXPY is therefore running at essentially the
full achievable bandwidth of the machine, and the load and store paths both show
4 sectors per request, which is exactly the ideal for a 4 byte access.

This is the anchor the whole suite depends on. If the simplest streaming kernel
had failed to approach the copy ceiling, the harness or the bandwidth
microbenchmark would have been wrong and nothing further would have been worth
reading. It did not, so the memory bound edge is trustworthy.

It is also where the theoretical byte count first misleads. At 4.2 million
elements SAXPY appears to reach \SI{1583}{GB/s}, more than twice the bus can
carry. Nothing is wrong with the timing: at that size the two vectors total
32\,MB and fit inside this card's 48\,MB L2, so the kernel is not touching DRAM
at all and the figure is L2 bandwidth wearing a DRAM label. Only once the working
set exceeds the cache do the numbers settle at the 560 to 566\,GB/s that DRAM can
actually sustain. This is precisely why every point on the roofline is labelled
with which byte count produced it.

\subsection{Transpose: coalescing, isolated}

The transpose pair does no floating point work, so nothing hides the memory
behaviour, and the counters give the cleanest result in the report:

\begin{center}
\begin{tabular}{lrr}
\toprule
 & naive & tiled \\
\midrule
load sectors per request & 4.00 & 4.00 \\
store sectors per request & 32.00 & 4.00 \\
duration (ms) & 0.163 & 0.086 \\
DRAM throughput (GB/s) & 143.5 & 350.2 \\
\bottomrule
\end{tabular}
\end{center}

A perfectly coalesced warp storing 4 byte values touches 128 contiguous bytes,
which is 4 sectors of 32 bytes. The naive kernel's stores take \emph{32} sectors
per request: because the transposed write strides by a whole row, every thread in
the warp lands in a different sector and the hardware issues eight times the
traffic it needs. Staging the tile through shared memory brings the stores back
to exactly 4 sectors per request, and the kernel runs 1.9 times faster.

The padded tile does its job too. The tiled transpose reports 40{,}532 shared
memory bank conflicts across the whole kernel, which is negligible next to the
tens of millions the unpadded GEMM tiles produced before I fixed them.

\subsection{The GEMM ladder is not memory bound at all}

The expectation going in was that the naive GEMM would be memory bound and that
shared memory tiling would fix it. The counters say otherwise, and the surprise
is the most interesting result in the project.

Naive and tiled GEMM at 2048 report L2 sector hit rates that are the same to
within measurement noise, and both move DRAM traffic of only 5 to 8\,GB/s against
a \SI{604}{GB/s} ceiling: between one and two percent of the memory system.
Neither kernel is close to memory bound. The naive kernel would move 68.7\,GB
from DRAM if every operand read went to memory; it actually moves 80.8\,MB, so
the cache absorbs better than 99.8 percent of its redundant loads.

That is why tiling does not pay here. Shared memory tiling exists to remove
redundant DRAM traffic, and on this card there is essentially none left to
remove: a 2048 square operand is 16\,MB and the working set fits inside the
48\,MB L2. What tiling still costs is a synchronisation per tile step and a
shared memory round trip, so it lands slightly \emph{behind} the naive kernel
rather than ahead of it. The classic optimisation is aimed at a bottleneck this
hardware no longer has at this problem size.

What does bind the naive kernel is instruction throughput: it reports 98.5
percent of peak sustained SM throughput while achieving under 2\,TFLOP/s of
useful arithmetic. It is saturating the SM issue and memory pipelines with an
enormous number of individually cheap loads that all hit cache. The fix is not to
move data less far, it is to ask for it fewer times, which is exactly what
register blocking does: each thread computes a 4 by 4 output tile from registers,
so one shared load feeds sixteen multiply adds instead of one. That rung reaches
\SI{5967}{GFLOP/s}, near four times the naive kernel, and the vectorized rung
reaches \SI{10325}{GFLOP/s}.

One caveat I want to state plainly: the naive GEMM's reported L2 hit rate is
100.39 percent. A hit rate cannot exceed 100 percent, and re-running the same
kernel gave 97.02, so this is noise in how the hardware attributes sector hits
rather than a real quantity. I report what the counter said and read it as
"essentially every access hits", rather than clamping the number to make the
table look tidy.

\subsection{A bug the counters found in my own kernels}

The bank conflict counter read 134.5 million for the register blocked GEMM and
93.4 million for the vectorized one, against 40 thousand for the padded
transpose. I had applied the \texttt{[TILE][TILE + 1]} padding to the transpose,
written a comment explaining why it matters, and then declared every GEMM shared
tile unpadded. The register blocked inner loop walks down a column of its tile,
so the address advances by a whole row per step; with an unpadded row length of
16 that stride is a multiple of the 32 bank width and the warp serialises on one
bank.

Padding the tiles is not quite free. The vectorized kernel reads its tile with a
\texttt{float4}, so row starts must remain 16 byte aligned; a one float pad would
make the row stride 260 bytes and turn an aligned vector load into undefined
behaviour. Padding by four floats preserves the alignment and still moves each
row off the colliding bank.

The measured outcome was not the clean sweep I expected. The vectorized kernel
gained 18 percent, from 8746 to \SI{10325}{GFLOP/s} at 4096. The register blocked
and tiled kernels did not move outside run to run noise. So bank conflicts were
the binding constraint in only one of the three, and I am reporting that rather
than implying the optimisation helped everywhere. Confirming the conflict counts
themselves dropped needs a further profiling pass.

\subsection{cuBLAS, and what the gap means}

cuBLAS reaches roughly \SI{22}{TFLOP/s}, about twice the best hand written rung
and near 68 percent of the theoretical compute ceiling. It is listed as a library
rather than as a rung of the ladder because the comparison is not like for like:
it selects among many tuned kernels per shape and uses instruction scheduling
that is not reachable from ordinary CUDA C. The honest reading is not that my
kernels are close, but that the ladder recovers a large fraction of the available
performance through three ideas (tiling for reuse, register blocking, and
vectorized access) and that the remaining factor of two is what a vendor library
buys.

\subsection{Thermals and clocks}

The NVML trace resolves one anomaly outright. The measured FP32 ceiling came out
at \SI{33.5}{TFLOP/s} against a theoretical \SI{32.26}{TFLOP/s}, which is 104
percent of a number that should be an upper bound. The clock trace shows the card
running at 2857\,MHz on average and 2865\,MHz at peak under load, while the CUDA
runtime reports a nominal 2625\,MHz. Recomputing the ceiling at the clock the card
actually ran gives \SI{35.2}{TFLOP/s}, and the microbenchmark then sits at 95
percent of it, which is where a good FMA loop belongs.

I have left the theoretical ceiling derived from the reported clock, because that
is the reproducible definition, and said here that the card exceeds it. Without
the power and clock trace this would have looked like a factor hiding somewhere
in the FLOP counting, and I would have gone looking for a bug that was never
there.
